The effect of clumpy outflows and disc anisotropy on quasar unification scenarios

### 2015-07-15 paper.tex
local laptop edits, intro and refs

Introduction focussing on key points

\begin{itemize}
\item standard wind + BALQSO introduction
\item focus on unification and that we are testing it
\item some discussion of scales, referencing e.g. reverb maps, variability, microlensing, Arav 
\item clumping background: stellar winds, clumping in AGN winds, variability
\end{itemize}
%%\item

\clearpage
Quasars and luminous active galactic nuclei (AGN) 
exhibit spectral energy distributions (SEDs) that
typically exhibit a series of strong emission lines (e.g. \la, \civ, \nv) 
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to 
emission from a geometrically thin, optically thick accretion disc 
surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.

In addition to the inflowing accreting material, 
outflowing material is ubiquitous in active galactic nuclei (AGN) 
and quasi-stellar objects (QSOs). These outflows can take the form of 
highly collimated radio jets \citep{bellonijet2010}, 
or mass-loaded `winds' emanating from the accretion disc. 
Approximately $20\%$ of QSOs exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BALs in quasar samples is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.

It is possible that the  diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.

As well as acting as a source of photoionized plasma, 
a disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN (REF).
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
In general, mass-loss treatments appear to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

\bibliography{mybib.bib}
\newpage
%\appendix
\end{document}

>>>>

% Introduction focussing on key points

% \begin{itemize}
% \item standard wind + BALQSO introduction
% \item focus on unification and that we are testing it
% \item some discussion of scales, referencing e.g. reverb maps, variability, microlensing, Arav 
% \item clumping background: stellar winds, clumping in AGN winds, variability
% \end{itemize}
%%\item 
Quasars and luminous active galactic nuclei (AGN) 
exhibit spectral energy distributions (SEDs) that
typically exhibit a series of strong emission lines (e.g. \la, \civ, \nv) 
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to 
emission from a geometrically thin, optically thick accretion disc 
surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars (REFs). These outflows can take the form of 
highly collimated radio jets \citep{bellonijet2010}(BETTER REF), 
or mass-loaded `winds' emanating from the accretion disc. 
Approximately $20\%$ of quasars exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.

It is possible that the  diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
% As well as acting as a source of photoionized plasma, a 
A disc wind may also have a profound effect on the structure and 
emergent continuum of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disc models.
In general, mass-loss treatments appear to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

Despite the clear importance of ADWs in understanding AGN SEDs and accretion physics, 
much of the underlying outflow physics remains highly uncertain. 
Several possible driving mechanisms for ADWs have been proposed, including
thermal pressure (REFs), magnetocentrifugal forces \citep{blandfordpayne} and 
radiation pressure on spectral lines \citep[`line-driving'][]{lucysolomon1970, MCGV95}.
Of these, line-driving is possibly the most attractive, due 
to the strong absorption lines seen in BALQSOs and the X-ray spectra of AGN (REFs).
The presence of line-locked features (REFs) and the so-called `ghost of \la' \citep{arav1996}
in the spectra of such objects also gives clearer evidence that line-driving is
at least partially contributing to the acceleration of the wind.
The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma. \cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.

An alternative solution is that the wind is clumped (REFs), possibly on multiple scale lengths.
Local density enhancements could lower the ionization parameter of the plasma
while still maintaining the same mass-loss rate and column density. 
An inhomogenous outflow has been proposed as a model for the BLR (REFs)
Clumping is expected in outflows and BLRs.

@@

\bibliography{mybib.bib,stellar.bib,h14.bib,h13.bib}
\newpage
%\appendix
\end{document}
